Foregrounds in Fourier Space for Redshifted 21~cm Observatories

### 2015-02-06 ms.tex
Complete summary section
from 3ed4ca5:
============

In ms.tex,

(*) Complete summary section
(*) Add emphasis on wide-field, pitchfork effects
(*) Add emphasis on effect of antenna shape on pitchfork signature

Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations. When the signal is expressed in units of temperature variance, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential. This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations with the MWA. By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more complex power spectrum estimators.

Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA, to the extent allowed by levels of uncertainty known in the foreground models and thermal noise fluctuations in measurements.

@@

As expected, diffuse emission in the primary lobe of the antenna power pattern is prominent on shorter antenna spacings. However, the most interesting result is its footprint on wide antenna spacings near the horizon delay limits --- an edge--heavy {\it two--pronged fork}--shaped signature. This is due to apparent shortening of wide antenna spacings in the direction of foreground emission far off--axis, thereby retaining their response to {\bf emission on large size scales}. On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}. Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified. In general, {\bf delay power spectrum} signatures of compact emission are center--heavy, in clear contrast to those from diffuse emission. A composite foreground model consisting of diffuse and compact foregrounds combines the characteristic individual shapes into a {\it pitchfork}--shaped structure in the {\it foreground wedge}. This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed through the MWA analysis pipelines. {\bf We conclude the {\it pitchfork} feature to be generic to all wide--field measurements, including EoR experiments}.

{\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied. In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and thus, preferable for EoR studies}.

>>>>

Our primary motivation in this work is to understand how the various bright foregrounds will manifest in three dimensional power spectrum of H{\sc i} from 21~cm reionization observations. {\bf In units of temperature variance}, the dynamic range between bright foregrounds and the 21~cm signal is expected to be $\sim 10^8$; a detailed understanding of how foregrounds can corrupt the 21~cm power spectrum is therefore essential. This analysis extends previous work by simulating the entire sky rather than just the central field of view and by providing a comparison with early observations with the MWA. By making use of the delay spectrum technique to estimate the power spectrum, we are able to observe the effects of foregrounds while avoiding entanglements with more {\bf sophisticated} power spectrum estimators.

{\bf We find that all wide--field instruments, typical of modern EoR observatories, imprint a characteristic {\it two--pronged fork} signature in delay spectra. There are two equivalent interpretations: delay bins near the horizon subtend larger solid angles and therefore contain larger integrated emission; or, foreshortening of baselines towards the horizon makes them sensitive to emission on large angular scales which match the inverse of their foreshortened lengths. These effects combined with higher sensitivity of antennas in the primary field of view results in a characteristic {\it pitchfork} signature. The amplitude of these generic signatures can be controlled by careful design of antenna aperture. In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination and thus preferable for EoR studies.}

Simulating in many important respects the response of the MWA to an all--sky foreground model that consists of diffuse Galactic emission from \citet{deo08} and bright point sources from the NVSS and SUMSS catalogs, we confirm that the modeled delay spectra are in agreement with data obtained with the MWA {\bf to within expected uncertainties in foreground models}.

@@

{\bf Diffuse emission in the primary field of view is prominent on shorter antenna spacings. However, it is also prominent near the horizon limits even on wide antenna spacings --- an effect of the wide--field nature of the measurement. On the other hand, compact emission predominantly maps onto central regions of the {\it foreground wedge}. Features arising from compact emission co--located with primary and secondary lobes of the antenna power pattern have been identified. In general, {\bf delay power spectrum} signatures of compact emission are center--heavy while those of diffuse emission are edge--heavy which results in the {\it pitchfork} signature. This will be distinctly visible when the thermal noise floor is sufficiently lowered, as longer observations are processed.}

% {\bf The role of antenna shape in determining the {\bf delay power spectrum} was also studied. In contrast to a dipole and a phased array such as an MWA tile, a dish such as the one proposed for HERA is found to yield the least foreground contamination both in the {\it foreground wedge} and the {\it EoR window} and thus, preferable for EoR studies}.
